Automated detections reveal the social information in the changing infant view

### 2020-08-14 writeup-paper/draft_apa_july2020.tex
latest paper draft

\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Using automated detections to quantify the effect of postural developments on the social information in the infant view},
            pdfkeywords={Postural developments, headcam, social information, face detection, pose
detection},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Using automated detections to quantify the effect of postural
developments on the social information in the infant view}

@@

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk. How do these
postural developments affect infants access to the social information
relevant for early learning? To address this question, we first created
a rich annotated dataset of naturalistic play sessions between infants
and their caregivers with sets of novel and familiar toys (N=36, 8-16
months of age), recording both egocentric and third-person video. We
then developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We found that
infants' posture and orientation to their caregiver changed dramatically
across this age range and modulated infants access to social
information; infants who were sitting or standing with their caregiver
at a close distance tended to have the most faces and hands in their
visual field. We also applied our automated method to a second
egocentric video dataset where one-year-olds explored a large play room
with their caregivers (Franchak et al., 2017), replicating and extending
their findings. Together, these results establish the validity of a new
method for detecting social information, confirm that motoric
developments play a significant role in shaping the social information
that infants have access to. We propose that the combined use of
head-mounted cameras and the application of new computer vision
techniques is a promising avenue for re-utilizing rich video datasets
collected across different labs and diverse contexts and understanding
the statistics of infants' visual and linguistic experience as they
change over development.
}

@@

Relatively little work, however, has quantified how often infants see
and use these kinds of social cues in naturalistic learing environments.
By using head-mounted cameras to record what infants see, researchers
have begun to document the infant egocentric perspective (Yoshida \&
Smith, 2008) and to quantify the information -- social and otherwise --
available to infants as they learn. While head-mounted camera data do
not provide explicit information on where infants are attending, some
work suggests that infants orient their head towards what they are
focusing on -- putting those people or objects in view (Yoshida \&
Smith, 2008). Initial recordings from in-lab play sessions revealed a
different view than many imagined: instead of being dominated by their
faces, the infant perspective contained close up views of primarily toys
and hands (Franchak, Kretch, Soska, \& Adolph, 2011; Yoshida \& Smith,
2008; Yu \& Smith, 2017). Subsequent research has revealed that the
infant view is far from a unitary construct, undergoing dramatic changes
as infants grow. Recordings from home environments suggest that the
viewpoints of very young infants -- less than 4 months of age -- do
indeed contain persistent and frequent views of faces (Fausey,
Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \& Smith, 2017) but that
the infant view tends to contain more and more hands as infants grow
older.

These changes in perspective are likely the downstream consequence of a
myriad factors, chief of which may be infants' evolving locomotive
abilities: an infants ability to sit, crawl, stand, or walk structures
the way they interact with the things and people in their world. These
changes can be rapid: the physical body of an infant can lengthen up to
2cm in 24 hours (Lampl, 1993), leading to changes in their motoric
abilities from one day to the next. These motoric developments can be
thought of as gateways that open up entirely new phases of development
(Iverson, 2010), and cause a cascade of changes in infants ability to
interact with their world and the people in it (Karasik, Tamis-LeMonda,
\& Adolph, 2014).

@@

To examine the robustness of our automated method and the
generalizability of our findings, we applied OpenPose (Cao et al., 2017)
to the dataset collected in Franchak et al. (2017), where one-year-olds
wore head-mounted eye-tracking cameras during a play session and their
in-the-moment posture was hand-annotated. Unlike the present dataset,
infants and caregivers roamed a large, open playroom and explored
different toys placed throughout. We thus analyzed this dataset with the
goal of both (1) validating our automated method on a very different
kind of dataset, attempting to replicate their primary findings
originally obtained with head-mounted eye-tracking (2) extending their
findings to ask whether whether the proportion of hands in view also
varies with infants' posture in an exploratory play session.

@@

A primary goal of this project was to provide a rich, open dataset for
other researchers to examine the effects of postural developments and
labeling behavior during naturalistic parent-child interactions. We thus
invited caregivers of 8, 12, and 16-month-olds to participate in play
sessions where they were provided with pairs of novel and familiar
objects (e.g., a ball and a microfiber duster, called a \enquote{zem})
in a small playroom in a lab (approximately 10' x 10'). Infants wore
head-mounted cameras equipped with a fish-eye lens (see Head-mounted
camera), a third-person camera captured a birds-eye view of the play
session. Using these video data, infants' posture and orientation to
their caregiver were hand-coded and annotated for the entirety of the
play session. All videos were transcribed, and we collected collected
MacArthur CDIs for all children who particpated. All materials have been
made publicly avaliable on Databrary for whom the parents provided
sharing consent (29/36 dyads) via
\url{https://nyu.databrary.org/volume/101}

@@

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below (see Table 1)/

@@

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam (e.g., failure to record, ran out of battery), 15 for
failing to wear the headcam, 10 for fewer than 4 minutes of headcam
footage, 5 for having multiple adults present, 5 for missing
Communicative Development Inventory (CDI) data, 2 for missing scene
camera footage, 1 for fussiness, and one for sample symmetry.
16-month-olds tolerated the head-mounted camera less well than younger
infants, leading to a higher exclusion rate. All inclusion decisions
were made independent of the results of subsequent analyses.

@@

\begin{figure}[H]
\includegraphics[width=0.4\linewidth]{images/view_angle_redone} \caption{Vertical field of view for two different headcam configurations (we used the lower in our current study).}\label{fig:headcam}
\end{figure}

We used a small, head-mounted camera (\enquote{headcam}) that was
constructed from a MD80 model camera attached to a soft elastic
headband. Videos captured by the headcam were 720x480 pixels with 25
frames per second.\footnote{Detailed instructions for creating this
  headcam can be found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, above).

Even with the fish-eye lens, the vertical field of view of the camera is
still considerably reduced compared to the child's field of view, which
spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
1988; Mayer, Fulton, \& Cummings, 1988). As we were originally primarily
interested in the presence of faces in the child's field of view, we
chose to orient the camera upwards to capture the entirety of the
child's upper visual field where the child is likely to see adult faces,
understanding that this decision limited our ability to detect hands
(especially those of the child, which are typically found at the bottom
of the visual field).

@@

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of familiar and novel objects. These pairs consisted of a ball
paired with a microfiber duster (a \enquote{zem}), a toy car paired with
a cheese grater (a \enquote{manu}), and a brush paired with a back
massager (a \enquote{tima}). Parents were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.} {[}\^{}3{]}:The first few participants played with a different
pair of objects (a toy cat and pedicure foam piece) that was replaced as
some infants chewed persistently on the foam.

@@

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture, M = 94.01s, SD = 225.18s; orientation; M = 94.01s, SD =
225.18s), and these rates did not vary substantially with the age of the
child. To assess the reliability of these annotations, a second coder
coded videos from three different children to calculate Cohen's kappa
(posture, irr = XX, orientation, irr = XX).

@@

One coder listened to all of the audio from the play sessions and marked
the exact timestamps whenever one of the novel or familiar objects was
named in any instance (e.g., \enquote{Look at the {[}ball{]}},
\enquote{Can you say {[}zem{]}?}); a second coder listened to the
majority of the play sessions (N=23 sessions) and also annotated all
naming events. To asesess reliability, we calculated the proportion of
naming events detected by the first coder that were also annotated by a
second coder within a sliding window. We found that 82.1\% of naming
events were detected within a 4 second window (+/- 2s), and 70.9\% of
namings events were detected within a 2 second window. We also obtained
full text transcriptions of the entire play sessions (with time stamps
marking 10s intervals). While these full transcriptions are not used in
the present analyses, they have been made publically avaliable for
future research.

@@

We evaluated three detection systems for the ability to measure infants'
access to faces. The first of these is the most commonly-used and widely
available face detection algorithm: Viola-Jones (Viola \& Jones, 2004).
We used this algorithm as a benchmark for performance, as while it can
achieve impressive accuracy in some situations, it is notoriously bad at
dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014).
We next tested the performance of two face detectors that both made use
of relatively recently developed Convolutional Neural Networks (CNNs) to
extract face information. The first algorithm was specifically optimized
for face detection, and the second algorithm was optimized to extract
pose information of all the individuals in an image, operationalized as
information about the position of 18 different body parts. For this
second algorithm (OpenPose; Cao et al., 2017), we used the agent's nose
(one of the body keypoints detected) to operationalize the presence of
faces, as any half of a face necessarily contains a nose.

@@

For face detection, we found that both OpenPose and MTCNN dramatically
outperformed ViolaJones, our baseline model, especially with respect to
the random sample, where ViolaJones generated a relatively high
proportion of both false negatives and false positives. When considering
only the composite F-score across all frames, MTCNN slightly
outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN and
OpenPose performed comparably with the random sample. Generally, MTCNN
exhibited higher precision, whereas OpenPose exhibited higher recall,
and these differences were most pronounced on the randomly sample
frames. In other words, while OpenPose generated slightly more false
positives than MTCNN, MTCNN missed several faces that were accurately
detected by OpenPose. When we restricted our analysis to high-confidence
detections from OpenPose (\textgreater{}.5 confidence; default threshold
for visualization), we found very high precision (P = 0.97), but much
lower recall (R = 0.64) and thus overall performance (F = 0.77),
indicating that these low-confidence detections often indexed actual
faces that were in the infant view. Figure \ref{fig:exampledetections}
shows example successful detections from OpenPose in each age group, and
Figure \ref{fig:failures} shows examples of missed faces as well as
false positives for context.

We next analyzed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.74). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showed low
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.37) and thus overall performance (F = 0.53).

@@

\subsubsection{Access to faces and
hands}\label{access-to-faces-and-hands}

First, we examined the proportion of face and hand detections as a
function of infants' age without considering their posture (see Figure
\ref{fig:detByAge}). While faces tended to be in the field-of-view
overall more often than hands, infants' head-mounted cameras were angled
slightly upward to capture the presence of faces, and hand detections
suffered from somewhat lower recall than face detections. We thus only
considered differences in the relative proportion of faces or hands in
view as a function of age, posture, and orientation, rather than
comparing them directly. Overall, we observed that 12-month-olds
appeared to have visual access to slightly fewer faces than 8 or
16-month-olds, creating a slight U-shaped function in face detections;
conversely, hand detections were showed a slight increase across this
age range, as reported in prior work (Fausey et al., 2016).

@@

However, these age-related trends were much smaller than the effect of
infant's postural developments on infants' visual access to faces and
hands. Infants' in-the-moment posture was a major factor both in how
many faces and hands were in view during the play session, as was their
orientation relative to their caregiver. This was quantified using two
generalized linear mixed-effect models estimating the proportion of
faces and hands that were in view, with orientation, posture, their
interaction, and scaled participant's age as fixed effects, and with
random slopes for infants' orientation and posture. A summary of the
coefficients of the models can be found in Tables 2 and 3.

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b = 0.99, SE = 0.07, Z = 14.54, P \textless{}
0.001; infant standing and CG close, b = 1.36, SE = 0.08, Z = 16.95, P
\textless{} 0.001) than infants who were lying down/crawling
(i.e.~prone). When caregivers were far away from their infants, face
detections were similarity higher (Face detections; infant sitting and
CG far, b = 0.75, SE = 0.07, Z = 10.48, P = P \textless{} 0.001; infant
standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = P \textless{}
0.001. Overall, we found that age did not remain a significant predictor
in accounting for the faces in view when modeling infants posture, their
orientation to their caregiver, and the interaction between these two
factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
= P = 0.697).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.31, SE = 0.08, Z = 3.72, P \textless{} 0.001; infant
standing and CG close, b=0.76, SE = 0.09, Z = 8.46, P \textless{} 0.001)
than infants who were lying down/crawling (i.e.~prone). Wrist detections
were highest when caregivers were far away from their infants and those
infants were standing (Wrist detections;infant standing and CG far, b =
1.76, SE = 0.11, Z = 15.87, P \textless{} 0.001). As with faces, age did
not remain a significant predictor in these models (Wrist detections;
Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.140).

>>>>

\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Automated detections reveal the social information in the changing infant view},
            pdfkeywords={Postural developments, headcam, social information, face detection, pose
detection},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Automated detections reveal the social information in the changing
infant view}

@@

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk. How do these
postural developments affect infants access to the social information
relevant for early learning? To address this question, we first created
a rich annotated dataset of naturalistic play sessions between infants
and their caregivers with sets of novel and familiar toys (N=36, 8-16
months of age), recording both egocentric and third-person video. We
then developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We found that
infants' posture and orientation to their caregiver changed dramatically
across this age range and modulated infants access to social
information; infants who were sitting or standing with their caregiver
at a close distance tended to have the most faces and hands in their
visual field. We also applied our automated method to a second
egocentric video dataset where one-year-olds explored a large play room
with their caregivers (Franchak et al., 2017), replicating and extending
their findings. Together, these results establish the validity of a new
method for detecting social information and confirm that motoric
developments play a significant role in shaping the social information
that infants have access to. We propose that the combined use of
head-mounted cameras and the application of new computer vision
techniques is a promising avenue for re-utilizing rich video datasets
and understanding how the statistics of infants' visual and linguistic
experience change over development.
}

@@

Relatively little work, however, has quantified how often infants see
and use these kinds of social cues in naturalistic learing environments.
By using head-mounted cameras to record what infants see, researchers
have begun to document the infant egocentric perspective (Yoshida \&
Smith, 2008) and to quantify the information -- social and otherwise --
available to infants as they learn. While head-mounted camera data do
not provide explicit information on where infants are attending, some
work suggests that infants orient their head towards what they are
focusing on -- putting those people or objects in view (Yoshida \&
Smith, 2008). Initial recordings using this technique during in-lab play
sessions revealed a different view than many imagined: instead of being
dominated by faces, the infant perspective contained close up views of
primarily toys and hands (Franchak, Kretch, Soska, \& Adolph, 2011;
Yoshida \& Smith, 2008; Yu \& Smith, 2017). Subsequent research has
revealed that the infant view is far from a unitary construct,
undergoing dramatic changes as infants grow. Recordings from home
environments suggest that the viewpoints of very young infants -- less
than 4 months of age -- do indeed contain persistent and frequent views
of faces (Fausey, Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \&
Smith, 2017) but that the infant view tends to contain more and more
hands as infants grow older.

These changes in perspective are likely the downstream consequence of a
myriad factors, chief of which may be infants' evolving locomotive
abilities: an infant's ability to sit, crawl, stand, or walk structures
the way they interact with the things and people in their world.
Infants' locomotive abilities can change rapidly: the physical body of
an infant can lengthen up to 2cm in 24 hours (Lampl, 1993), leading to
changes in their physical abilities from one day to the next. These
motoric developments can be thought of as gateways that open up entirely
new phases of development (Iverson, 2010), and cause a cascade of
changes in infants ability to interact with their world and the people
in it (Karasik, Tamis-LeMonda, \& Adolph, 2014).

@@

To examine the robustness of our automated method and the
generalizability of our findings, we applied OpenPose (Cao et al., 2017)
to the dataset collected in Franchak et al. (2017), where one-year-olds
wore head-mounted eye-tracking cameras during a play session and their
in-the-moment posture was hand-annotated. Unlike the present dataset,
infants and caregivers roamed a large, open playroom and explored
different toys placed throughout. We thus analyzed this dataset with the
goal of both (1) validating our automated method on a very different
kind of dataset, attempting to replicate their primary findings
originally obtained with a head-mounted eye-tracker (2) examining
whether the proportion of hands in view also varies with infants'
posture in an exploratory play session.

@@

A primary goal of this project was to provide a annotated, open dataset
for other researchers to examine the effects of postural developments
and labeling behavior during naturalistic parent-child interactions. We
thus invited caregivers of 8, 12, and 16-month-olds to participate in
play sessions where they were provided with pairs of novel and familiar
objects (e.g., a ball and a microfiber duster, called a \enquote{zem})
in a small playroom in a lab (approximately 10' x 10'). Infants wore
head-mounted cameras equipped with a fish-eye lens (see Head-mounted
camera), a tripod-mounted camera captured a third-person view of the
play session. Using these video data, infants' posture and orientation
to their caregiver were hand-coded and annotated for the entirety of the
play session. All videos were transcribed, and MacArthur CDIs were
collected for all children who particpated. All materials have been made
publicly avaliable on Databrary for whom the parents provided sharing
consent (29/36 dyads) via \url{https://nyu.databrary.org/volume/101}

@@

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below (see Table 1).

@@

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam (e.g., failure to record, ran out of battery), 15 for
failing to wear the headcam, 10 for fewer than 4 minutes of headcam
footage, 5 for having multiple adults present, 5 for missing
Communicative Development Inventory (CDI) data, 2 for missing scene
camera footage, 1 for fussiness, and 1 for sample symmetry.
16-month-olds tolerated the head-mounted camera less well than younger
infants, leading to a higher exclusion rate. All inclusion decisions
were made independent of the results of subsequent analyses.

@@

We used a head-mounted camera (\enquote{headcam}) that was constructed
from a small camera attached to a soft elastic headband.\footnote{Detailed
  instructions for creating this headcam can be found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
Initial participants wore an MD-80 camera, which was then replaced by a
Veho pro camera which had better battery life and a larger view angle.
The view angle of the MD-80 camera was \(32^{\circ}\) horizontal by
\(24^{\circ}\) vertical, and we attached a fish-eye lens to the camera
to increase the view angle to \(64^{\circ}\) horizontal by
\(46^{\circ}\) vertical. The view angle of the Veho pro camera was
wider, \(47^{\circ}\) horizontal by \(36^{\circ}\) vertical. Videos
captured by MD-80/Veho cameras were 640x480/720x480 pixels,
respectively, and both cameras had a frame rate of \textasciitilde{}30
frames per second. To ensure taht detections across these different
cameras were comparable, we excluded detections from the outer edges of
the videos taken with the MD-80 cameras based on these view angle
differences in our main analyses (i.e.~excluding top/bottom 13\% of the
frames, and left/right 10\% of the frames).

However, the vertical field of view of the cameras were still
considerably reduced compared to the infants' field of view, which spans
around 100--120\(^{\circ}\) in the vertical dimension by 6-7 months of
age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton, 1988; Mayer,
Fulton, \& Cummings, 1988). As we were primarily interested in the
presence of faces in the child's field of view, we chose to orient the
camera upwards to capture the entirety of the child's upper visual field
where the child is likely to see adult faces, understanding that this
decision limited our ability to detect hands (especially those of the
child, which are typically found at the bottom of the visual field).

@@

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of familiar and novel objects. These pairs consisted of a ball
paired with a microfiber duster (a \enquote{zem}), a toy car paired with
a cheese grater (a \enquote{manu}), and a brush paired with a back
massager (a \enquote{tima}). Parents were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.}

@@

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture, M = 104.99s, SD = 233.86s; orientation; M = 104.99s, SD =
233.86s), and these rates did not vary substantially with the age of the
child. To assess the reliability of these annotations, a second coder
annotated videos from three different children to calculate Cohen's
kappa (posture, irr = .82; caregiver orientation, irr = .73).

@@

One coder listened to all of the audio from the play sessions and marked
the exact timestamps whenever one of the novel or familiar objects was
named in any instance (e.g., \enquote{Look at the {[}ball{]}},
\enquote{Can you say {[}zem{]}?}); a second coder listened to the
majority of the play sessions (N = 23 sessions) and also annotated all
naming events. To asesess reliability, we calculated the proportion of
naming events detected by the first coder that were also annotated by
the second coder within a sliding window. We found that 82.1\% of naming
events were detected within a 4 second window (+/- 2s), and 70.9\% of
namings events were detected within a 2 second window (+/- 1s). We also
obtained full text transcriptions of the entire play sessions (with time
stamps marking 10s intervals). While these full transcriptions are not
used in the present analyses, they have been made publically avaliable
for future research.

@@

We evaluated three detection systems for the ability to measure infants'
visual access to faces. The first of these is the most commonly-used and
widely available face detection algorithm: Viola-Jones (Viola \& Jones,
2004). We used this algorithm as a benchmark for performance, as while
it can achieve impressive accuracy in some situations, it is notoriously
bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox,
2014). We next tested the performance of two face detectors that both
made use of relatively recently developed Convolutional Neural Networks
(CNNs) to extract face information. The first algorithm was specifically
optimized for face detection, and the second algorithm was optimized to
extract pose information of all the individuals in an image,
operationalized as information about the position of 18 different body
parts. For this second algorithm (OpenPose; Cao et al., 2017), we used
the agent's nose (one of the body keypoints detected) to operationalize
the presence of faces, as any half of a face necessarily contains a
nose.

@@

For face detection, we found that both OpenPose and MTCNN dramatically
outperformed ViolaJones, our baseline model, especially with respect to
the random sample, where ViolaJones missed many face that were in view.
When considering only the composite F-score across all frames, MTCNN
slightly outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN
and OpenPose performed comparably with the random sample. Generally,
MTCNN exhibited higher precision, whereas OpenPose exhibited higher
recall, and these differences were most pronounced on the randomly
sample frames. In other words, while OpenPose generated slightly more
false positives than MTCNN, MTCNN missed several faces that were
accurately detected by OpenPose. When we restricted our analysis to
high-confidence detections from OpenPose (\textgreater{}.5 confidence;
default threshold for visualization), we found very high precision (P =
0.97), but much lower recall (R = 0.64) and thus overall performance (F
= 0.77), indicating that these low-confidence detections often indexed
actual faces that were in the infant view. Figure
\ref{fig:exampledetections} shows example successful detections from
OpenPose in each age group, and Figure \ref{fig:failures} shows examples
of missed faces as well as false positives for context.

We next assessed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.73). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showed low
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.36) and thus overall performance (F = 0.52).

@@

\subsubsection{Access to faces and hands across
age}\label{access-to-faces-and-hands-across-age}

First, we examined the proportion of face and hand detections as a
function of infants' age without considering their posture (see Figure
\ref{fig:detByAge}). While faces tended to be in the field-of-view
overall more often than hands, infants' head-mounted cameras were angled
slightly upward to capture the presence of faces, and hand detections
suffered from somewhat lower recall than face detections. We thus only
considered differences in the relative proportion of faces or hands in
view as a function of age, posture, and orientation, rather than
comparing them directly. Overall, we did not observe strong age related
trends from 8-16 months of age; if anything, face detections showed a
slight u-shaped pattern, with 12-month-olds having slightly fewer faces
in their visual field that 8- or 16-month-olds.

@@

In contrast, infant's locomotor developments had a major effect on the
faces and hands that were in the field of view. Infants' in-the-moment
posture was a major factor both in how many faces and hands were in view
during the play session, as was their orientation relative to their
caregiver (see Figure \ref{fig:detByPosOrient}). This was quantified
using two generalized linear mixed-effect models estimating the
proportion of faces and hands that were in view, with orientation,
posture, their interaction, and scaled participant's age as fixed
effects, and with random slopes for infants' orientation and posture. A
summary of the coefficients of the models can be found in Tables 2 and
3.

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b = 0.90, SE = 0.07, Z = 13.80, P \textless{}
0.001; infant standing and CG close, b = 1.23, SE = 0.08, Z = 15.44, P
\textless{} 0.001) than infants who were lying down/crawling
(i.e.~prone). When caregivers were far away from their infants, face
detections were similarity higher (Face detections; infant sitting and
CG far, b = 0.63, SE = 0.07, Z = 9.16, P \textless{} 0.001; infant
standing and CG far, b = 1.23, SE = 0.09, Z = 14.43,P \textless{}
0.001). Infants' age was not significant predictor in accounting for the
faces in view (Face detections; Age (scaled), b = 0.04, SE = 0.11, Z =
0.42, P = 0.671).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.27, SE = 0.08, Z = 3.35, P \textless{} 0.001; infant
standing and CG close, b=0.53, SE = 0.09, Z = 6.07, P \textless{} 0.001)
than infants who were lying down/crawling (i.e.~prone). Wrist detections
were highest when caregivers were far away from their infants and those
infants were standing (Wrist detections; infant standing and CG far, b =
1.57, SE = 0.11, Z = 14.16, P \textless{} 0.001). As with faces, age was
not a significant predictor in these models (Wrist detections; Age
(scaled), b = 0.09, SE = 0.11, Z = 0.86, P = 0.390).
